Predicting Diamond Pricing: A Generalized Linear Model Analysis

### 2024-12-04 paper/paper.qmd
Remove the typo in the paper then upload

---
title: "Predicting Diamond Pricing: A Generalized Linear Model Analysis"
subtitle: "Carat Weight Dominates While Cut Plays a Smaller Role in Determining Price"
author: 
  - Yawen Tan
thanks: "Code and data are available at: [https://github.com/YawennnnnnTan/Analysis-of-Diamond-Price]"
date: today
date-format: long
abstract: "This paper investigates how the Four Cs—carat weight, color, clarity, and cut—affect diamond pricing and develops a predictive model using a Generalized Linear Model (GLM) with a Gamma distribution. The analysis shows that carat weight has the strongest positive influence on price, followed by clarity and color, while cut has a relatively smaller impact. The model demonstrates robust performance, providing a reliable framework for forecasting diamond prices and understanding market patterns. These findings enhance pricing transparency, help the diamond industry adapt to shifting consumer preferences, and offer valuable insights into the economic and cultural factors shaping diamond valuation."
format: pdf
number-sections: true
bibliography: 
  - references .bib
toc: true
---

@@

Diamonds have long symbolized luxury, romance, and prestige, with their valuation primarily determined by the Four Cs: Carat weight, Color, Clarity, and Cut. The Gemological Institute of America established this standard to assess diamond quality consistently worldwide. Each of these attributes uniquely influences a diamond's market value. Carat size measures the diamond's mass, with larger diamonds being rarer and typically more valuable. Color grading evaluates the absence of color, ranging from D (colorless) to Z (light yellow or brown), with colorless diamonds commanding higher prices. Clarity shows internal and external imperfections, where fewer inclusions and blemishes enhance value. Cut quality, particularly in round brilliant diamonds, affects light reflection and overall brilliance, significantly impacting aesthetic appeal and price (@GIA4Cs). While the 4Cs provide a framework for valuation, accurately predicting diamond prices remains complex due to the interplay of these characteristics and market dynamics. Recent advancements in data analytics and machine learning have enabled more precise predictions of diamond prices based on their features. For instance, research of @Basha2024 published in the International Journal of System Assurance Engineering and Management, evaluates 23 machine learning models, including regression techniques, to predict diamond prices. The study shows the effectiveness of models like CatBoost Regressor and XGBoost Regressor in achieving high accuracy after hyperparameter tuning. Similarly, @Sharma2021 also conducts a comparative analysis of multiple supervised machine learning models, identifying Random Forest Regressor as particularly effective in predicting diamond prices. Despite these advancements, there remains a need for analyses diamond price based on other models to enhance the accuracy and reliability of diamond price predictions. This research fills this gap by employing a Generalized Linear Model.

@@

The barplot on the left of @fig-color shows the distribution of diamond color from D to K. The barplot shows that the number of diamonds with color grades E and F is the largest, indicating that these nearly colorless diamonds are more popular in the market, probably because they have high quality and cost performance. In contrast, the number of completely colorless D-class diamonds and K-class diamonds with slight yellow is small, which may be due to their rarity and color deviation, respectively, resulting in low demand. This picture directly reflects the supply of diamonds of different color grades in the market and their potential demand trends.
The violin chart on the right of @fig-color shows the distribution of diamond prices in different color grades. The violio chart shows that the prices of diamonds with color grades D to F are generally higher, especially those with color grades D and E, which shows a wider price distribution range, indicating that the prices of high-end diamonds with these color grades can increase significantly. In contrast, the price distribution of diamonds with color grades I to K is more concentrated and the overall price is lower, reflecting that the price fluctuation of diamonds close to yellow is less in the market. In addition, there are significant outliers in top colors such as D and E, which may be due to the unusually high price of diamonds with large carats or other high-quality characteristics. This figure clearly reveals the influence of color grade on price and its market value distribution characteristics.

@@

# 绘制小提琴图
violin_plot_color <- ggplot(cleaned_data, aes(x = color, y = price)) +
  geom_violin(trim = FALSE, fill = "lightblue", alpha = 0.7) +  # 添加小提琴图
  geom_boxplot(width = 0.2, outlier.color = "salmon", outlier.shape = 8, alpha = 0.7) +  # 在小提琴图中叠加箱线图
  labs(
    title = "Violin Plot of Price by Color",
    x = "Diamond Color",
    y = "Diamond Price (USD)"
  ) +
  theme_minimal()+
  theme(plot.title = element_text(size = 12, face = "bold",hjust = 0.5 ))

@@

The diamond carat size, a continuous numerical variable representing the weight of the diamond, ranges from approximately 0.30 to 2.02 carats in the dataset. One carat is equivalent to 200 milligrams, making it a key factor in determining the diamond's size and price.
The histogram on the left of @fig-carat_size shows the distribution of the Carat Size of diamonds. On the whole, it presents a right-skewed distribution, indicating that the number of small carats (such as 0.5 to 1.0 carats) is the largest in the market, while the number of larger carats is gradually decreasing. The red dotted line in the figure indicates the average carat size, and it can be seen that the carat size of most diamonds is concentrated on the left side of the average, which further reflects the dominant position of small carat diamonds in the market. This reflects that consumers have higher demand for small carats, while large carats are scarce because of their rarity and high price.
The scatter chart on the right of @fig-carat_size shows the relationship between the carat size of diamonds and the price. It can be clearly seen that the price rises rapidly with the increase of carat size, showing a nonlinear growth trend. Especially in the range of close to 1.0 carats and larger carats, the price increase is more significant. This shows that the carat size has an important influence on the diamond price, but when the Dancla size exceeds a certain value (such as 1.5 carats), the price growth rate tends to be flat. In addition, the fitted black curve further clearly shows this nonlinear relationship, indicating that the marginal effect of carat size on price decreases.

@@

@fig-correlation_heat shows several key insights about the relationships between diamond attributes and price. Carat size shows a strong positive correlation with price, indicating that it is one of the primary factors influencing diamond value—larger diamonds tend to be more expensive. In contrast, cut (e.g., cutExcellent, cutVery Good) demonstrates a weaker correlation with price, suggesting that cut grade alone has a limited direct impact on diamond cost but may work in conjunction with other factors such as carat size or color. Similarly, color (e.g., colorD to colorK) shows a weak correlation with price, with diamonds closer to colorless (e.g., colorD and colorE) potentially commanding higher prices, but the overall effect is minimal. Clarity (e.g., clarityIF to claritySI2) also exhibit a modest relationship with price, reflecting their role as a contributing but less dominant factor in determining diamond value. Overall, carat size emerges as the most influential attribute, while other variables like cut, color, and clarity play supporting roles.
@fig-correlation_heat also presents the relationships between the predictor variables, showing that the correlations among them are generally weak. For example, attributes like cut quality, clarity, color, and carat size do not exhibit strong intercorrelations, as most of the corresponding cells are closer to white or light blue. This suggests that the predictor variables are relatively independent, reducing the risk of multicollinearity in the analysis. The weak correlations among these variables ensure that each contributes uniquely to the model, providing a robust foundation for predicting diamond price.

@@

one_hot_encoded <- cbind(
  correlated_data %>% select(price, carat_size), # 数值变量
  model.matrix(~ color + clarity + cut - 1, data = correlated_data, 
               contrasts.arg = list(
                 color = contrasts(correlated_data$color, contrasts = FALSE),
                 clarity = contrasts(correlated_data$clarity, contrasts = FALSE),
                 cut = contrasts(correlated_data$cut, contrasts = FALSE)
               ))
)

@@

\begin{align}
\log(\mu_i) &= \beta_0 + \beta_1 x_{1i} + \sum_{j} \beta_{2j} \cdot \text{color}_{ij} \nonumber \\
&\quad + \sum_{k} \beta_{3k} \cdot \text{cut}_{ik} + \sum_{l} \beta_{4l} \cdot \text{clarity}_{il}.
\end{align}

@@

\begin{align}
\text{price} &= \log(\text{price}) \cdot \sigma + \mu.
\end{align}

@@

# Create the plot
ggplot(model_summary_df, aes(x = Estimate, y = reorder(Predictors, Estimate))) +
  geom_point(color = "palevioletred") +  # Coefficients as points
  geom_errorbarh(aes(xmin = CI_Lower, xmax = CI_Upper), height = 0.2) +  # Confidence intervals
  geom_text(aes(label = round(Estimate, 3)), hjust = -0.3, size = 3) +  # Lable estimates
  theme_minimal() +  # Minimal theme
  labs(
    x = "Coefficient Estimate",
    y = "Predictors"
  ) + 
  theme(axis.text.y = element_text(size = 8))  # Adjust y-axis text size

@@

@@

## Exploration of Model Result {#sec-exploration}

@@

Carat size has the strongest impact on price. Larger diamonds are often associated with wealth, success, and love, making them highly desirable for significant events like engagements and weddings. This focus on size shows how people value physical attributes that are easily noticeable, often linking material possessions to social status and personal achievements. Economically, the preference for size drives demand for larger diamonds, encouraging industries to focus on producing and marketing them as rare and valuable.

@@

Although the data set is valuable, there are several limitations, which affect the universality and depth of the research results. A key limitation is the limited price range, which only includes diamonds with a price between $1,000 and $10,000. This does not include low-priced and high-end luxury diamonds, which may miss the unique pricing model of these market segments. In addition, the price distribution is seriously tilted to the right, and most of the prices are concentrated between 2500 and 5000 dollars. This imbalance may lead to deviation in the analysis, because diamonds with higher prices may be affected by different market factors and their representation is insufficient.

Another limitation is that the data set narrowly focuses on physical properties, such as carat size, color, clarity and cutting, while ignoring background factors such as brand reputation, certification sources and market trends. These external factors will significantly affect pricing, but they are not taken into account in the data. In addition, the data set lacks geographical and cultural details, which can provide insights into regional differences in preferences and needs. This omission limits the global applicability of the research results, because diamond pricing often varies with the market and consumer base.

The dataset also simplifies some predictor variables, particularly cut quality, by classifying them into broad groups (Ideal, Excellent, Very Good, and Good) without accounting for finer details like proportions or symmetry. These details, such as the arrangement of facets or crown angles, are crucial for evaluating a diamond's brilliance and could provide a deeper explanation for price variations. Similarly, the color and clarity variables are not fully represented. For instance, diamond color is graded from D (colorless) to Z (light yellow or brown), but the dataset only includes colors from D to K, leaving out lower-grade colors that could provide additional insights. Moreover, the absence of time-related factors limits the ability to analyze how diamond prices change over time due to shifts in consumer preferences, market trends, or seasonal effects.

@@

```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.pos="H"}
#| label: fig-prediction_actual
#| fig-cap: "Actual vs Predicted Values"
test_data <- read_parquet(here("data/02-analysis_data/diamondprice_test_data.parquet"))

test_data$predicted_price <- predict(model_glm, newdata = test_data, type = "response")

ggplot(data = test_data, aes(x = 1:nrow(test_data))) +
  geom_line(aes(y = predicted_price), color = "lightblue", size = 1, linetype = "solid") + 
  geom_point(aes(y = price), color = "palevioletred", size = 2) + 
  labs(
    x = "Index",
    y = "Price"
  ) +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5)) +
  scale_y_continuous(breaks = NULL)

@@

@fig-metrics shows how the Generalized Linear Model (GLM) was evaluated on the test dataset using three key metrics: Mean Squared Error (MSE), Mean Absolute Error (MAE), and Root Mean Squared Error (RMSE). The MSE of 1.627 indicates that the overall squared differences between the predicted and actual diamond prices are relatively small, suggesting that the model captures the relationship between the predictors and the response variable reasonably well. However, since MSE amplifies the impact of larger errors, it is complemented by RMSE for better interpret ability.The MAE of 0.517 shows that, on average, the absolute difference between the predicted and actual values is low. This highlights the model's strong ability to provide accurate predictions in most cases and reflects that its performance is not heavily skewed by outliers. Compared to MSE, MAE is less sensitive to extreme errors, offering a more balanced perspective on prediction accuracy. The RMSE of 1.276, being the square root of MSE, provides a direct measure of the standard deviation of prediction errors in the original data's units. While RMSE tends to be slightly higher than MAE because it gives more weight to larger errors, its relatively low value confirms the model's stability and reliable performance on the test data. However, the gap between RMSE and MAE suggests the presence of some larger errors or outliers that may require further investigation.Overall, these metrics indicate that the GLM performs well in predicting diamond prices, with reasonable accuracy and controlled error levels.

@@

The "Residuals vs. Leverage" plot highlights a few potentially influential points (e.g., 1930, 1899, 1924) with higher leverage. Although none of these points exceed Cook’s distance thresholds, their presence warrants closer examination to ensure they do not disproportionately affect the model.

@@

par(mfrow = c(2, 2))
plot(model_glm)

@@

The questionnaire design includes:
-   Closed-form questions (multiple choice, Likert scale) to evaluate consumer preferences.
-   Open-ended questions gain insight into the factors affecting prices.
The questionnaire was optimized by small-scale pretest, focusing on reducing the ambiguity of questions and the fatigue of respondents

>>>>

---
title: "Predicting Diamond Pricing: A Generalized Linear Model Analysis"
subtitle: "Carat Weight Dominates While Cut Plays a Smaller Role in Determining Price"
author: 
  - Yawen Tan
thanks: "Code and data are available at: [https://github.com/YawennnnnnTan/Analysis-of-Diamond-Price]"
date: today
date-format: long
abstract: "This paper investigates how the Four Cs—carat weight, color, clarity, and cut—affect diamond pricing and develops a predictive model using a Generalized Linear Model (GLM) with a Gamma distribution. The analysis shows that carat weight has the strongest positive influence on price, followed by clarity and color, while cut has a relatively smaller impact. The model demonstrates robust performance, providing a reliable framework for forecasting diamond prices and understanding market patterns. These findings enhance pricing transparency, help the diamond industry adapt to shifting consumer preferences, and provide opinions into the economic and cultural factors shaping diamond valuation."
format: pdf
number-sections: true
bibliography: 
  - references .bib
toc: true
---

@@

Diamonds have long symbolized luxury, romance, and prestige, with their valuation primarily determined by the Four Cs: Carat weight, Color, Clarity, and Cut. The Gemological Institute of America established this standard to assess diamond quality consistently worldwide. Each of these attributes uniquely influences a diamond's market value. Carat size measures the diamond's mass, with larger diamonds being rarer and typically more expensive. Color grading evaluates the absence of color, ranging from D (colorless) to Z (light yellow or brown), with colorless diamonds commanding higher prices. Clarity shows internal and external imperfections, where fewer inclusions and blemishes enhance value. Cut quality, particularly in round brilliant diamonds, affects light reflection and overall brilliance, significantly impacting aesthetic appeal and price (@GIA4Cs). While the 4Cs provide a framework for valuation, accurately predicting diamond prices remains complex due to the interplay of these characteristics and market dynamics. Recent advancements in data analytics and machine learning have enabled more precise predictions of diamond prices based on their features. For instance, research of @Basha2024 published in the International Journal of System Assurance Engineering and Management, evaluates 23 machine learning models, including regression techniques, to predict diamond prices. The study shows the effectiveness of models like CatBoost Regressor and XGBoost Regressor in achieving high accuracy after hyperparameter tuning. Similarly, @Sharma2021 also conducts a comparative analysis of multiple supervised machine learning models, identifying Random Forest Regressor as particularly effective in predicting diamond prices. Despite these advancements, there remains a need for analyses diamond price based on other models to enhance the accuracy and reliability of diamond price predictions. This research employs a Generalized Linear Model to analyze more.

@@

The barplot on the left of @fig-color shows the distribution of diamond color from D to K. The barplot shows that the number of diamonds with color grades E and F is the largest, indicating that these nearly colorless diamonds are more popular in the market, probably because they have high quality and cost performance. In contrast, the number of completely colorless D-class diamonds and K-class diamonds with slight yellow is small, which may be due to their rarity and color deviation, respectively, resulting in low demand. This picture directly reflects the supply of diamonds of different color grades in the market and their potential demand trends.
The violin chart on the right of @fig-color shows the distribution of diamond prices in different color grades. The violio chart shows that the prices of diamonds with color grades D to F are generally higher, especially those with color grades D and E, which shows a wider price distribution range, indicating that the prices of high-end diamonds with these color grades can increase significantly. In contrast, the price distribution of diamonds with color grades I to K is more concentrated and the overall price is lower, reflecting that the price fluctuation of diamonds close to yellow is less in the market. In addition, there are significant outliers in top colors such as D and E, which may be due to the unusually high price of diamonds with large carats or other high-quality characteristics. This figure clearly shows the influence of color grade on price and its market value distribution characteristics.

@@

violin_plot_color <- ggplot(cleaned_data, aes(x = color, y = price)) +
  geom_violin(trim = FALSE, fill = "lightblue", alpha = 0.7) + 
  geom_boxplot(width = 0.2, outlier.color = "salmon", outlier.shape = 8, alpha = 0.7) + 
  labs(
    title = "Violin Plot of Price by Color",
    x = "Diamond Color",
    y = "Diamond Price (USD)"
  ) +
  theme_minimal()+
  theme(plot.title = element_text(size = 12, face = "bold",hjust = 0.5 ))

@@

The diamond carat size, a continuous numerical variable representing the weight of the diamond, ranges from approximately 0.30 to 2.02 carats in the dataset. One carat is equivalent to 200 milligrams, making it a key factor in determining the diamond's size and price.
The histogram on the left of @fig-carat_size shows the distribution of the Carat Size of diamonds. On the whole, it presents a right-skewed distribution, indicating that the number of small carats (such as 0.5 to 1.0 carats) is the largest in the market, while the number of larger carats is gradually decreasing. The red dotted line in the figure indicates the average carat size, and it can be seen that the carat size of most diamonds is concentrated on the left side of the average, which further reflects the dominant position of small carat diamonds in the market. This reflects that consumers have higher demand for small carats, while large carats are scarce because of their rarity and high price.
The scatter chart on the right of @fig-carat_size shows the relationship between the carat size of diamonds and the price. It can be clearly seen that the price rises quickly with the increase of carat size, showing a nonlinear growth trend. Especially in the range of close to 1.0 carats and larger carats, the price increase is more significant. This shows that the carat size has an important influence on the diamond price, but when the Dancla size exceeds a certain value (such as 1.5 carats), the price growth rate tends to be flat. In addition, the fitted black curve further clearly shows this nonlinear relationship, indicating that the marginal effect of carat size on price decreases.

@@

@fig-correlation_heat shows several key points about the relationships between diamond attributes and price. Carat size shows a strong positive correlation with price, indicating that it is one of the primary factors influencing diamond value—larger diamonds tend to be more expensive. In contrast, cut (e.g., cutExcellent, cutVery Good) demonstrates a weaker correlation with price, suggesting that cut grade alone has a limited direct impact on diamond cost but may work in conjunction with other factors such as carat size or color. Similarly, color (e.g., colorD to colorK) shows a weak correlation with price, with diamonds closer to colorless (e.g., colorD and colorE) potentially commanding higher prices, but the overall effect is minimal. Clarity (e.g., clarityIF to claritySI2) also exhibit a modest relationship with price, reflecting their role as a contributing but less dominant factor in determining diamond value. Overall, carat size emerges as the most influential attribute, while other variables like cut, color, and clarity play supporting roles.
@fig-correlation_heat also presents the relationships between the predictor variables, showing that the correlations among them are generally weak. For example, attributes like cut quality, clarity, color, and carat size do not exhibit strong intercorrelations, as most of the corresponding cells are closer to white or light blue. This suggests that the predictor variables are relatively independent, reducing the risk of multicollinearity in the analysis. The weak correlations among these variables ensure that each contributes uniquely to the model, providing a robust foundation for predicting diamond price.

@@

one_hot_encoded <- cbind(
  correlated_data %>% select(price, carat_size), 
  model.matrix(~ color + clarity + cut - 1, data = correlated_data, 
               contrasts.arg = list(
                 color = contrasts(correlated_data$color, contrasts = FALSE),
                 clarity = contrasts(correlated_data$clarity, contrasts = FALSE),
                 cut = contrasts(correlated_data$cut, contrasts = FALSE)
               ))
)

@@

\begin{align}
\log(\mu_i) &= \upsilon_0 + \upsilon_1 x_{1} + \upsilon_2 x_{2}+\upsilon_3 x_{3}+... 
\end{align}

- $\beta_0$: Intercept term.
- $\upsilon_i$: Coefficients of predictors.
- $x_i$: Predictors.

@@

\begin{align}
\text{price} &= \log(\text{standardized\ and\ exponentiated\ price}) \cdot \sigma + \mu.
\end{align}

@@

# Create the plot
ggplot(model_summary_df, aes(x = Estimate, y = reorder(Predictors, Estimate))) +
  geom_point(color = "palevioletred") +  # Coefficients as points
  geom_errorbarh(aes(xmin = CI_Lower, xmax = CI_Upper), height = 0.2) +  # Confidence intervals
  geom_text(aes(label = round(Estimate, 2)), hjust = -0.3, size = 3) +  # Lable estimates
  theme_minimal() +  # Minimal theme
  labs(
    x = "Coefficient Estimate",
    y = "Predictors"
  ) + 
  theme(axis.text.y = element_text(size = 8))  # Adjust y-axis text size

@@

To further explain the result of the model, we will show an example of diamond price predictions.

-   Carat size: 1.2
-   Color: D
-   Cut: Good
-   Clarity: VVS2

Based on above features of diamond, we firstly calculate the log(standardized and exponential price). From @fig-modelresults, we can know intercept and the coefficients each variables:

-   Intercpet: -4.81 
-   carat size: 3.74
-   color D: 1.34
-   cut Good: 0, because it is the reference level
-   clarity VVS2: 0.89

We plog these value into formula (3) and can get the log(standardized and exponential price) is around 1.91. Then we will calculate the actual price of the diamond. We plog 1.91 into formula (4) and get the price is around 8678.07. Therefore, we know a diamond with 1.2 carat size, color D, good cut and VVS2 clarity should be approximately 8678.07 USD.

@@

## Discussion of Model Result {#sec-exploration}

@@

Carat size has the strongest impact on price. Larger diamonds are often associated with wealth, success, and love, making them highly desirable for significant events like engagements and weddings. This focus on size shows how people value physical attributes that are easily noticeable, often linking material possessions to social status and personal achievements. Economically, the preference for size drives demand for larger diamonds, encouraging industries to focus on producing and marketing them as rare and 
expensive.

@@

Although the data set is good, there are several limitations, which affect the universality and depth of the research results. A key limitation is the limited price range, which only includes diamonds with a price between 1,000 USD and 10,000 USD. This does not include low-priced and high-end luxury diamonds, which may miss the unique pricing model of these market segments. In addition, the price distribution is seriously tilted to the right, and most of the prices are concentrated between 2500 and 5000 dollars. This imbalance may lead to deviation in the analysis, because diamonds with higher prices may be affected by different market factors and their representation is insufficient.

Another limitation is that the data set narrowly focuses on physical properties, such as carat size, color, clarity and cutting, while ignoring background factors such as brand reputation, certification sources and market trends. These external factors will significantly affect pricing, but they are not considered in the data. In addition, the data set lacks geographical and cultural details, which can provide opinions into regional differences in preferences and needs. This omission limits the global applicability of the research results, because diamond pricing often varies with the market and consumer base.

The dataset also simplifies some predictor variables, particularly cut quality, by classifying them into broad groups (Ideal, Excellent, Very Good, and Good) without accounting for finer details like proportions or symmetry. These details, such as the arrangement of facets or crown angles, are crucial for evaluating a diamond's brilliance and could provide a deeper explanation for price variations. Similarly, the color and clarity variables are not fully represented. For instance, diamond color is graded from D (colorless) to Z (light yellow or brown), but the dataset only includes colors from D to K, leaving out lower-grade colors that could provide additional ideas. Moreover, the absence of time-related factors limits the ability to analyze how diamond prices change over time due to shifts in consumer preferences, market trends, or seasonal effects.

@@

```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.pos="H"}
#| label: fig-prediction_actual
#| fig-cap: "Actual vs Predicted Values"

# Load and deal with the dataset
test_data <- read_parquet(here("data/02-analysis_data/diamondprice_test_data.parquet"))
test_data$price <- scale(test_data$price, center = TRUE, scale = TRUE)
test_data$price<-exp(test_data$price)
test_data$color <- factor(test_data$color)
test_data$cut <- factor(test_data$cut)
test_data$clarity <- factor(test_data$clarity)

## Generate predicted prices using the generalized linear model (model_glm)
test_data$predicted_price <- predict(model_glm, newdata = test_data, type = "response")

# Plot the actual vs. predicted prices
ggplot(data = test_data, aes(x = 1:nrow(test_data))) +
  # Add a line for predicted prices
  geom_line(aes(y = predicted_price), color = "lightblue", size = 1, linetype = "solid") +
  # Add points for actual prices
  geom_point(aes(y = price), color = "palevioletred", size = 2) + 
  labs(
    x = "Index",
    y = "Price"
  ) +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5)) +
  scale_y_continuous(breaks = NULL)

@@

@fig-metrics shows how the Generalized Linear Model (GLM) was evaluated on the test dataset using three key metrics: Mean Squared Error (MSE), Mean Absolute Error (MAE), and Root Mean Squared Error (RMSE). The MSE of 1.627 indicates that the overall squared differences between the predicted and actual diamond prices are relatively small, suggesting that the model captures the relationship between the predictors and the response variable reasonably well. However, since MSE amplifies the impact of larger errors, it is complemented by RMSE for better interpret ability.The MAE of 0.517 shows that, on average, the absolute difference between the predicted and actual values is low. This highlights the model's strong ability to provide accurate predictions in most cases and reflects that its performance is not heavily skewed by outliers. Compared to MSE, MAE is less sensitive to extreme errors, providing a more balanced perspective on prediction accuracy. The RMSE of 1.276, being the square root of MSE, provides a direct measure of the standard deviation of prediction errors in the original data's units. While RMSE tends to be slightly higher than MAE because it gives more weight to larger errors, its relatively low value confirms the model's stability and reliable performance on the test data. However, the gap between RMSE and MAE suggests the presence of some larger errors or outliers that may require further investigation.Overall, these metrics indicate that the GLM performs well in predicting diamond prices, with reasonable accuracy and controlled error levels.

@@

The last plot highlights a few potentially influential points (e.g., 1930, 1899, 1924) with higher value. Although none of these points exceed Cook’s distance thresholds, their presence warrants closer examination to ensure they do not disproportionately affect the model.

@@

# Set the plotting layout to a 2x2 grid for displaying four plots on one screen
par(mfrow = c(2, 2))

# Generate diagnostic plots for the generalized linear model (model_glm)
# 1. Residuals vs Fitted: Checks for patterns in residuals to assess linearity and homoscedasticity
# 2. Normal Q-Q Plot: Assesses whether residuals follow a normal distribution
# 3. Scale-Location Plot: Examines homogeneity of variance (spread of residuals)
# 4. Residuals vs Leverage: Identifies influential data points
plot(model_glm)

@@

The questionnaire design includes:
-   Closed-form questions (multiple choice, Likert scale) to evaluate consumer preferences.
-   Open-ended questions gain opinions into the factors affecting prices.
The questionnaire was optimized by small-scale pretest, focusing on reducing the ambiguity of questions and the fatigue of respondents
